\textsf{Are large language models superhuman chemists?}

### 2024-10-21 src/tex/ms.tex
Update src/tex/ms.tex
Co-authored-by: Martiño Rios-Garcia <147555961+MrtinoRG@users.noreply.github.com>

To compile our benchmark corpus, we utilized a broad list of sources (see \Cref{sec:curation}), ranging from completely novel manually crafted questions, over university exams to semi-automatically generated questions based on curated subsets of data in chemical databases.
For quality assurance, all questions have been reviewed by at least one scientist in addition to the original curator and automated checks.
Importantly, our large pool of questions encompasses a wide range of topics and question types (\Cref{fig:corpus}). The topics range from general chemistry to more specialized fields such as inorganic, analytical or technical chemistry.
We also classify the questions based on what skills are required to answer them. Here, we distinguish between questions that require knowledge, reasoning, calculation, intuition or a combination of these.
Moreover, to allow for a more nuanced evaluation of the models capabilities, the questions are also classified by difficulty by the annotator.

@@

One interesting finding of recent research is that foundation models can judge interestingness.\autocite{zhang2024omniopenendednessmodelshuman, Argyle_2023}
If models could do so for chemical compounds, this would open opportunities for novel optimization approaches.
Such open-ended tasks, however, depend on an external observer defining what interestingness is.\autocite{hughes2024openendednessessentialartificialsuperhuman}
Here, we posed models the same question~\textcite{Choung_2023} asked chemists at a drug company: \enquote{Which of the two compounds do you prefer?} (in the context of an early virtual screening campaign setting).
While chemists showed a fair degree of interrater agreement, most of our models do not correlate with the preferences of expert chemists---even if they perform well on many other tasks in \chembench (\Cref{tab:performance_table_human_subset})
This indicates that using preference tuning for chemical settings is a promising approach to explore in future research.

>>>>

To compile our benchmark corpus, we utilized a broad list of sources (see \Cref{sec:curation}), ranging from completely novel manually crafted questions, over university exams to semi-automatically generated questions based on curated subsets of data in chemical databases.
For quality assurance, all questions have been reviewed by at least two scientists in addition to the original curator and automated checks.
Importantly, our large pool of questions encompasses a wide range of topics and question types (\Cref{fig:corpus}). The topics range from general chemistry to more specialized fields such as inorganic, analytical or technical chemistry.
We also classify the questions based on what skills are required to answer them. Here, we distinguish between questions that require knowledge, reasoning, calculation, intuition or a combination of these.
Moreover, to allow for a more nuanced evaluation of the models capabilities, the questions are also classified by difficulty by the annotator.

@@

One interesting finding of recent research is that foundation models can judge interestingness.\autocite{zhang2024omniopenendednessmodelshuman, Argyle_2023}
If models could do so for chemical compounds, this would open opportunities for novel optimization approaches.
Such open-ended tasks, however, depend on an external observer defining what interestingness is.\autocite{hughes2024openendednessessentialartificialsuperhuman}
Here, we posed models the same question~\textcite{Choung_2023} asked chemists at a drug company: \enquote{Which of the two compounds do you prefer?} (in the context of an early virtual screening campaign setting) (see \Cref{tab:chembench_corpus_cognitive}).
While chemists showed a fair degree of interrater agreement, most of our models do not correlate with the preferences of expert chemists---even if they perform well on many other tasks in \chembench (\Cref{tab:performance_table_human_subset})
This indicates that using preference tuning for chemical settings is a promising approach to explore in future research.
